---
title: "Motor Trend Analysis"
author: "Ricardo Arango Giraldo"
date: "20 de junio de 2018"
output:
  pdf_document: default
  html_document: default
---
## Executive Summary

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

- Is an automatic or manual transmission better for MPG.

- Quantify the MPG difference between automatic and manual transmissions.

## Load Data

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models). It consists of 32 observations on 11 variables.

```{r,echo=TRUE}
library(datasets)
data(mtcars)
names(mtcars)
```

## Is an automatic or manual transmission better for MPG?

For automatic:
```{r,echo=TRUE}
summary(mtcars[mtcars$am==0,])
```



For manual:
```{r,echo=TRUE}
summary(mtcars[mtcars$am==1,])
```

Hence, the mean of mpg is greater for manual (at 24.4) than automatic (at 17.1).

Investigating further..

## Quantify the MPG difference between automatic and manual transmissions.

```{r,echo=TRUE}
boxplot(mpg ~ am, data = mtcars, xlab = "Transmission", ylab = "Miles per gallon", main="Miles per gallon by Transmission Type")
```

Manual (represented by 1) has a higher mean for mpg than automatic (represented by 0).

## Hypothesis Testing

```{r,echo=TRUE}
aggregate(mpg~am, data = mtcars, mean)
```

The mean transmission for manual is 7.24mpg higher than automatic. Let alpha=0.5.
```{r,echo=TRUE}
auto <- mtcars[mtcars$am == 0,]
manual <- mtcars[mtcars$am == 1,]
t.test(auto$mpg, manual$mpg)
```

Since p-value = 0.001374, we reject the null hypothesis. There is a major difference between mpg of manual and automatic transmissions.

```{r,echo=TRUE}
m<-lm(mpg~am,data=mtcars)
summary(m)
```

From the above, we may conclude that automatic run at 17.15mpg, while manual have 7.24 more mpg.

Also, R^2 is 0.36, hence the model only accounts for 36% variance.

Performing multivariate linear regression:
```{r,echo=TRUE}
model <- lm(mpg~am + wt + hp + cyl, data = mtcars)
anova(m,model)
```

The final model is below:
```{r,echo=TRUE}
summary(model)
```

## Conclusion

This model explains 84.9% of the variance. It may be concluded that on average, manual transmissions have 1.478 more mpg than automatic.

## APPENDIX

```{r,echo=TRUE}
plot(model)
```

Hence, the residuals are normally distributed, and homoskedastic.